Estimating Learning Dynamics in the Routing Game

### 2015-10-15 paper/paper.tex
Progress commit

\documentclass{sig-alternate-ipsn13}
\usepackage{ntheorem}
\usepackage{algorithm}
\usepackage[noend]{algorithmic}
\usepackage{todonotes}
\usepackage{enumitem}

@@

@@

We conclude this section by observing that while we chose to apply the model to a simple prediction task, the estimated model can be used, more generally, in any a receding-horizon optimal control problem, by using the current estimate of the model as a plant in the control problem.
%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
\begin{figure*}
\centering
\includegraphics[width=.75\textwidth]{images/user_interface}
%\missingfigure[figwidth=.3\textwidth]{System block diagram}
\caption{\footnotesize Screenshot of the client side of the routing game application. The table is the main interface, and can be used by the player to set weights on the different paths, using the sliders. The weights determine the flow distribution $x^{(t+1)}_k$. The table also show the previous flows ($x^{(t)}_k$), the previous costs $(\ell^{(t)}_k)$, and the cumulative costs $L^{(t)}_k = \sum_{\tau \leq t} \ell^{(\tau)}_k$. Clicking a path will also highlight the path on the graph. The bottom charts show the full history of flows $(x^{(\tau)}_k)_{\tau \leq t}$, costs $(\ell^{(\tau)}_k)_{\tau \leq t}$, and cumulative losses $(L^{(\tau)})_{\tau \leq t}$.}
\label{fig:game_interface}
\label{fig:system_block_diagram}
\end{figure*}
%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
%============================================================================================
\section{The Routing Game Web Application}
\label{sec:experiment}

@@

\begin{figure}[h]
  \centering
  \missingfigure{System block diagram to be added later}
  \caption{\footnotesize System block diagram to be added later.}
  \label{fig:system_block_diagram}
\end{figure}

@@

% \tikzstyle{block} = [draw, fill=white, rectangle,
% minimum height=3em, minimum width=6em]

% \tikzstyle{block} = [draw,fill=blue!20,minimum size=2em]

% \tikzstyle{sum} = [draw, fill=white, circle, node distance=1cm]
% \tikzstyle{input} = [coordinate]
% \tikzstyle{output} = [coordinate]
% \tikzstyle{pinstyle} = [pin edge={to-,thin,black}]

% \def\radius{.7mm}
% \tikzstyle{branch}=[fill,shape=circle,minimum size=3pt,inner sep=0pt]

% \begin{tikzpicture}[auto, node distance=2cm,>=latex']

%   \node [input, name=input] {};
%   \node [sum, right of=input] (sum) {};
%   % \node [block, right of=sum] (controller) {Controller};
%   \node [block, right of=sum, node distance=3cm] (system) {Network};

%   \node [output, right of=system] (output) {};
%   % \node [block, below of=system] (measurements) {Measurements};

%   % \draw [draw,->] (input) -- node {$\bar x(t)$} (sum);
%   \draw [->] (sum) -- node {$\bar x(t)$} (system);
%   % \draw [->] (y) |- (measurements);
%   % \draw [->] (measurements) -| node[pos=0.99] {$-$}
%   % node [near end] {$y_m$} (sum);

%   \node at (0,-1) (input1) {$i_1$};
%   \node[block, above left = of system, anchor=west] at (2,-1) (block1) {$p_1$};
%   \draw[->] (block1.east) -- (sum);
%   \draw[->] (system) -- (block1);

%   \node at (0,-2) (input2) {$i_2$};
%   \node[block, left of=system, anchor=west] at (2,-2) (block2) {$p_2$};
%   \draw[->] (block2.east) -- (sum);
%   \draw[->] (system) -- (block2);

%   \node at (0,-3) (input3) {$i_3$};
%   \node[block, below left = of system, anchor=west] at (2,-3) (block3) {$p_3$};
%   \draw[->] (block3.east) -- (sum);
%   \draw[->] (system) -- (block3);

%   % % Draw blocks, inputs and outputs
%   % \foreach \y in {1,2,3} {
%   %   \node at (0,-\y) (input\y) {$i_\y$};
%   %   \node[block, left of=system, above of =system] at (2,-\y) (block\y) {$p_\y$};
%   %   \draw[->] (input\y) -- (block\y);
%   %   \draw[->] (block\y.east) -- (sum);
%   %   \draw[->] (system) -- (block\y);
%   % }

%   % Calculate branch point coordinate
%   \path (input1) -- coordinate (branch) (block1);

%   % Define a style for shifting a coordinate upwards
%   % Note the curly brackets around the coordinate.
%   \tikzstyle{s}=[shift={(0mm,\radius)}]
%   % It would be natural to use the yshift or xshift option, but that does
%   % not seem to work when shifting coordinates.

%   \draw[->] (branch) node[branch] {}{ % draw branch junction
%     \foreach \c in {2} {
%       % Draw semicircle junction to indicate that the lines are
%       % not connected. The intersection between the lines are
%       % calculated using the convenient -| syntax. Since we want
%       % the semicircle to have its center where the lines intersect,
%       % we have to shift the intersection coordinate using the 's'
%       % style to account for this.
%       [shift only] -- ([s]input\c -| branch) arc(90:-90:\radius)
%       % Note the use of the [shift only] option. It is not necessary,
%       % but I have used it to ensure that the semicircles have the
%       % same size regardless of scaling.
%     }
%   };

% \end{tikzpicture}

@@

First, we solve Problem~\eqref{eq:estimation_eta} to estimate the learning rate sequence one term at a time. Then, we use the parameterized form $\eta^{(t)}_k = \eta_k^{(0)} t^{-\alpha_k}$, and estimate $\eta_k^{(0)}$ and $\alpha_k^{(0)}$ by solving the problem~\ref{eq:estimation_alpha_eta0}. The results of both methods are shown in Figure~\eqref{fig:moving_average_learning_rate}
When we estimate one term at a time, the resulting sequences have very large variations, and to better visualize them, we plot a moving average (over a window of $3$ iterations).

@@

%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
\begin{figure}[h!]
  \centering
  \includegraphics[width=0.37\textwidth]{images/moving_average_learning_rate}
  \includegraphics[width=0.37\textwidth]{images/sequence_rate}
  \caption{\footnotesize Estimated sequences of learning rates in logarithmic scale. In the top figure, we estimate one term of the sequence at a time then plot a moving average with a window length equal to $3$. In the bottom figure, we estimate for each player the initial term $\eta^{(0)}_k$ and the decay rate $\alpha_k$.}
  \label{fig:moving_average_learning_rate}
  \label{fig:sequence_rate}
\end{figure}
%-----------------------------------------------------------------------------------------------------------------------------------------------------------------

>>>>

\documentclass{sig-alternate-ipsn13}
\usepackage{ntheorem}
\usepackage{algorithm}
\usepackage[noend]{algorithmic}
\usepackage{todonotes}
\usepackage{enumitem}
\usepackage{pgfplots}
\usepackage{filecontents}

@@

\tikzstyle{block} = [draw, rectangle, minimum height=2em, minimum width=4em]
%fill=blue!20
\tikzstyle{sum} = [draw, fill=blue!20, circle, node distance=1cm]
\tikzstyle{input} = [coordinate] \tikzstyle{output} = [coordinate]
\tikzstyle{pinstyle} = [pin edge={to-,thin,black}]

@@

We conclude this section by observing that while we chose to apply the model to a simple prediction task, the estimated model can be used, more generally, in any a receding-horizon optimal control problem, by using the current estimate of the model as a plant in the control problem.
%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
\begin{figure*}
\centering
\includegraphics[width=.80\textwidth]{images/user_interface.pdf}
%\missingfigure[figwidth=.3\textwidth]{System block diagram}
\caption{\footnotesize Screenshot of the client side of the routing game application. The table is the main interface, and can be used by the player to set weights on the different paths, using the sliders. The weights determine the flow distribution $x^{(t+1)}_k$. The table also show the previous flows ($x^{(t)}_k$), the previous costs $(\ell^{(t)}_k)$, and the cumulative costs $L^{(t)}_k = \sum_{\tau \leq t} \ell^{(\tau)}_k$. Clicking a path will also highlight the path on the graph. The bottom charts show the full history of flows $(x^{(\tau)}_k)_{\tau \leq t}$, costs $(\ell^{(\tau)}_k)_{\tau \leq t}$, and cumulative losses $(L^{(\tau)})_{\tau \leq t}$.}
\label{fig:game_interface}
\label{fig:system_block_diagram}
\end{figure*}
%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
%============================================================================================
\section{The Routing Game Web Application}
\label{sec:experiment}

@@

%\begin{figure*}[h]
%  \centering
%  \missingfigure{System block diagram to be added later}
%  \caption{\footnotesize System block diagram to be added later.}
%  \label{fig:system_block_diagram}
%\end{figure*}

%
% \tikzstyle{block} = [draw, fill=white, rectangle,
% minimum height=3em, minimum width=6em]
%
% \tikzstyle{block} = [draw,fill=blue!20,minimum size=2em]
%
% \tikzstyle{sum} = [draw, fill=white, circle, node distance=1cm]
% \tikzstyle{input} = [coordinate]
% \tikzstyle{output} = [coordinate]
% \tikzstyle{pinstyle} = [pin edge={to-,thin,black}]

 \def\radius{.7mm}
 \tikzstyle{branch}=[fill,shape=circle,minimum size=3pt]

\tikzstyle{dotted}= [dash pattern=on .01mm off 2.5mm,
                                         line cap=round]

\begin{figure}[h]
 \begin{tikzpicture}[auto, node distance=2cm,>=latex']

    \node[block] (system) at (6, -2) {Server};
    \node[output] at (7.3, -2) (sys_output) {};
    \node[sum, left of=system] (sum)  at (4, -2) {};
    \node[block, above left = of sum, anchor=west, align = left] at (1,-1) (block1) {$p_1$};
    \node[block, left = of sum, anchor=west, align = left] at (1.6,-2) (block2) {$p_2$};
    \node[block, below left = of sum, anchor=west, align = left] at (1,-3) (block3) {$p_3$};

   \coordinate[below = 0.3cm of block1] (b1_turn);
   \coordinate[above = 0.3cm of block2] (b2_turn);
   \coordinate[above = 0.3cm of block3] (b3_turn);
   \coordinate(system_b1_turn) at (3.7, 1.55);
   \coordinate(system_b2_turn) at (3.7, -0.85);
   \coordinate(system_b3_turn) at (3.7, -3.265);

   \draw[->] (sum) -- node {$\bar x^{(t)}$} (system);
   \draw[-] (system) -- (sys_output);

   \draw[->] (block1.east) -| node [near start] {$\bar x_1^{(t)} $} (sum) ;
   \draw[->] (sys_output) |- (system_b1_turn) -| node [near start, above right = 0.0cm and 1.5cm of block1] {$(\bar \ell_{1}^{(t)}, \bar \ell^{(t-1)}_{1}, \dots)$} (block1.north);

   \draw[->] (block2) |- node [near end]  {$\bar x_2^{(t)} $} (sum);
   \draw[->] (sys_output) |- (system_b2_turn) -| node [near start, above right = 0.0cm and 1.5cm of block1] {$(\bar \ell_{2}^{(t)}, \bar \ell_2^{(t-1)}, \dots)$}  (block2.north);

   \draw[->] (block3) -| node [near start] {$\bar x_3^{(t)} $}  (sum);
   \draw[->] (sys_output) |- (system_b3_turn) -| node [near start, above right = 0.0cm and 1.5cm of block1] {$(\bar \ell_{3}^{(t)}, \bar \ell_3^{(t-1)}, \dots)$}  (block3.north);

\draw[dotted,line width = 2mm] (0.3,-5.1) -- (0.3, -5.8);

  \end{tikzpicture}
  \label{fig:system_diagram}
  \caption{Block diagram of the system. At iteration $t$, each player $p_k$ uses the cost vector history $(\bar \ell_{k}^{(t-1)}, \bar \ell_k^{(t-2)}, \dots)$ to make the routing decision vector $\bar x_k^{(t)}$. The server then computes the cumulative flow on each edge and returns the local cost vector history $(\bar \ell_{k}^{(t)}, \bar \ell_k^{(t-1)}, \dots)$ for player $p_k$, then the game proceeds to iteration $t+1$.}
 \end{figure}

@@

@@

First, we solve Problem~\eqref{eq:estimation_eta} to estimate the learning rate sequence one term at a time. Then, we use the parameterized form $\eta^{(t)}_k = \eta_k^{(0)} t^{-\alpha_k}$, and estimate $\eta_k^{(0)}$ and $\alpha_k^{(0)}$ by solving the problem~\ref{eq:estimation_alpha_eta0}. The results of both methods are shown in Figure~\eqref{fig:moving_average_learning_rate}
When we estimate one term at a time, the resulting sequences have very large variations, and to better visualize them, we plot a moving average (over a window of $5$ iterations).

@@

%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
\begin{figure}[h!]
  \centering
  \includegraphics[width=0.37\textwidth]{images/moving_average_learning_rate}
  \includegraphics[width=0.37\textwidth]{images/sequence_rate}
  \caption{\footnotesize Estimated sequences of learning rates in logarithmic scale. In the top figure, we estimate one term of the sequence at a time then plot a moving average with a window length equal to $5$. In the bottom figure, we estimate for each player the initial term $\eta^{(0)}_k$ and the decay rate $\alpha_k$.}
  \label{fig:moving_average_learning_rate}
  \label{fig:sequence_rate}
\end{figure}
%-----------------------------------------------------------------------------------------------------------------------------------------------------------------
